Increasing Productivity Amid Stable Nutrient Regimes in Rhode Island Lakes and Reservoirs

### 2019-05-24 manuscript/manuscript.Rmd
first round of writing done...  Lot's to fill in but my co-authors can do that...

In addition to the annualized trends we address with this study, there are other trends that may be of interest.  First, what are the trends for water quality at finer temporal scales such as monthly or seasonal trends?  Anecdotal evidence in Rhode Island [SHOULD REFERENCE THIS SOMEHOW] points to greater increases in temperature earlier and later in the growing season and suggests a lengthening of the growing season. Furthermore, preliminary analysis of the URIWW data back this up with mean temperature for May 1993 to May 1995 cooler by nearly a degree than mean temperature for May 2014 through May 2016.  Second, are trends influenced by the current state of a given waterbody?  For instance, are oligotrophic lakes showing stronger trends than eutorphic lakes or are all lakes showing similar trends regardless of current trohpic status?  These questions are beyond the scope of this study, but all warrant further, careful investigation.

@@

- Use of site specific z-scores
    - Other approaches available, but this one is useful because it provided readily interpretable results, robust to variations in sampling stations (i.e. added later in period, removed, etc.), and based off of approach with long history of use in exploring long-term trends [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface]

- Why this is possible - Fantastic WW data

>>>>

In addition to the annualized trends we address with this study, there are other trends that may be of interest.  First, what are the trends for water quality at finer temporal scales such as monthly or seasonal trends?  Anecdotal evidence in Rhode Island [SHOULD REFERENCE THIS SOMEHOW] points to greater increases in temperature earlier and later in the growing season and suggests a lengthening of the growing season. Furthermore, preliminary analysis of the URIWW data back this up with mean temperature for May 1993 to May 1995 cooler by nearly a degree than mean temperature for May 2014 through May 2016.  Second, are trends influenced by the current state of a given waterbody?  For instance, are oligotrophic lakes showing stronger trends than eutorphic lakes or are all lakes showing similar trends regardless of current trophic status?  These questions are beyond the scope of this study, but all warrant further, careful investigation.

@@

The analysis approach we used here, site-specific z-scores, is not a novel method, but using it to examine water quality trends is novel as we could find few examples of using it specifically for water quality trends [NEED TO DO THIS LIT SEARCH].  This approach, and in particular non-scaled anomalies, does have a long history in the analysis of trends in climate [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface].  We simply built on this methods and adapted it for use with long-term water quality trends.  And while other methods are valid and robust (e.g. [@Oliver_2017]), we chose averaged site-specific z-scores as they, in our opinion, provide readily interpretable results, especially for communicating to less technical audiences.  In addition, the site-specific z-scores are robust to variations in sampling effort and in the timing of inclusion of given sampling locations (i.e. added later in a time period or removed).  Lastly, this analysis is only possible because of the availability of sound, long-term data on water quality in Rhode Island.  In other words, without the URIWW data and the commitment and participation of nearly XXX volunteers our analysis would have been impossible.  Going forward, it is important to appreciate the role that volunteer monitoring and citizen science programs can play in the analysis of long term environmental trends.
